---
title: "Task 4: Halloween Candy Data 🍬"
output: html_notebook
---

#**INTRO** 

This data looks at different peoples opinions for candy that could be revived at Halloween across three years. 


#**CLEANING PROCESS🧹** 

I started the cleaning with one data frame at a time, lucky the 3 data frames had and some similar issues.

* **Column names:** The Raw data had a lot of issue with columns which had nothing to do with candy, this meant that having to filter out any columns that I found irrelevant to the task. *'e.g "Which day do you prefer, Friday or Sunday?"'* 
I also had to standardize a lot of columns over the three data frames. *'e.g your_gender = gender'*
             
* **Age:** The age column in 2015 was as a character string with inputs that needed to cleaned, I made it as an integer column and rounded and .something ages to the nearest whole number, anything that was text input that didn't make sense was changed to NA.
I also then also filter out any one under the age 2 and over 120.  

* **Country:** This columns was a mess with lots of different varieties for the same country. I used a pattern *'e.g uk_pattern <- "UK|ENGLAND|UNITED K+|U.K|SCOTLAND|IRELAND"'* to standardize this column into six categories (USA, UK, EU, CANADA, ASIA, OTHER)

* **Year:** As I knew we had to combine the data frames I made sure each data frame had a year column before the join so we don't loses the input year of the data.

* **Pivot:** To make the data more readable I used a pivot_longer so the data would be easier to do analysis on. Making 'candy'
and 'rating' columns. 

<br>

#**LET'S ANALYSE🔎** 


```{r}
library(tidyverse)
library(here)
```


```{r}
candy_all_data <- read_csv(here("clean_data/candy_all_data.csv"))
```

#**AGE**


**What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)** 

```{r}
candy_all_data %>% 
  filter(!is.na(rating)) %>% 
  summarise(rating = n())
```
Answer: 576925

<br>


**What was the average age of people who are going out trick or treating?**

```{r}
candy_all_data %>% 
  filter(going_trick_or_treating == "Yes") %>% 
  summarise(age = mean(age)) %>% 
  round()

```

Answer: rounded 39

<br>

**What was the average age of people who are not going trick or treating?**


```{r}
candy_all_data %>% 
  filter(going_trick_or_treating == "No") %>% 
  summarise(age = mean(age)) %>% 
  round()

```
Answer: rounded 39

<br>

#**MOST RATINGS**

**For each of joy, despair and meh, which candy bar received the most of these ratings?**

*'JOY:'*

```{r}
candy_all_data %>% 
  select(candy, rating) %>% 
  filter(rating == "JOY") %>% 
  group_by(candy) %>% 
  summarise(rating = n())  %>% 
  arrange(desc(rating)) %>% 
  head(5)

```
For joy: any_full_sized_candy_bar comes out top with reese_s_peanut_butter_cup in close second. 

<br>

*'FOR MEH:'*
```{r}
candy_all_data %>% 
  select(candy, rating) %>% 
  filter(rating == "MEH") %>% 
  group_by(candy) %>% 
  summarise(rating = n()) %>% 
  arrange(desc(rating)) %>% 
  head(5)

```
For Meh: lollipops got the most ratings 

<br>

*'FOR DESPAIR:'*

```{r}
candy_all_data %>% 
  select(candy, rating) %>% 
  filter(rating == "DESPAIR") %>% 
  group_by(candy) %>% 
  summarise(rating = n()) %>% 
  arrange(desc(rating)) %>% 
  head(5)

```
For Despair: gum form baseball cards came out on top 

<br>

#**STARBURST**

**How many people rated Starburst as despair?**

```{r}
candy_all_data %>% 
  filter(candy == "starburst") %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(candy = n())
```

ANSWER: 1866

<br>

#**RANKING SYSTEM**

***'For the next three questions, count despair as -1, joy as +1, and meh as 0'***

```{r}

rating_candy_data <- candy_all_data %>%
  mutate(number_rating = case_when(
    str_detect(rating, pattern = "DESPAIR") ~ -1,
    str_detect(rating, pattern = "JOY") ~ 1,
    str_detect(rating, pattern = "MEH") ~ 0,
  ))

```



**What was the most popular candy bar by this rating system for each gender in the dataset ?**

<br>

*'Starting with male:'*
```{r}
rating_candy_data %>%
  select(gender, candy, number_rating, rating) %>%
  filter(gender == "Male") %>%
  group_by(candy) %>%
  summarise(male_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(male_count)) %>%
  head(5)

```
The answer give the top answer being any_full_sized_candy_bar but as this a very vague answer I would say that reese_s_peanut_butter_cups get top place.

<br>

*'Now female:"* 

```{r}
rating_candy_data %>%
  select(gender, candy, number_rating, rating) %>%
  filter(gender == "Female") %>%
  group_by(candy) %>%
  summarise(female_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(female_count)) %>%
  head(5)
```
same as the male excluding any_full_sized_candy_bar the top is reese_s_peanut_butter_cups

<br>

*'others:'*

```{r}
rating_candy_data %>%
  select(gender, candy, number_rating, rating) %>%
  filter(gender == "Other") %>%
  group_by(candy) %>%
  summarise(other_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(other_count)) %>%
  head(5)
```
Like the men and women the top answer is full_sized_candy_bar but excluding this the top answer is kit_kat 


#**YEAR**

<br>

**What was the most popular candy bar in each year?**

*'2015'*
```{r}
rating_candy_data %>%
  select(year, candy, number_rating, rating) %>%
  filter(year == "2015") %>%
  group_by(candy) %>%
  summarise(year_2015_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(year_2015_count)) %>%
  head(5)

```
Answer: any_full_sized_candy_bar with reeses second 

<br>

*'2016'*
```{r}
rating_candy_data %>%
  select(year, candy, number_rating, rating) %>%
  filter(year == "2016") %>%
  group_by(candy) %>%
  summarise(year_2016_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(year_2016_count)) %>%
  head(5)
```
Answer: any_full_sized_candy_bar with kit_kat second 

<br>

*'2017'*

```{r}
rating_candy_data %>%
  select(year, candy, number_rating, rating) %>%
  filter(year == "2017") %>%
  group_by(candy) %>%
  summarise(year_2017_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(year_2017_count)) %>%
  head(5)
```
Answer: any_full_sized_candy_bar with reeses second 

<br>

#**COUNTRY**


**What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?**

*'USA'*
```{r}
rating_candy_data %>%
  select(country, candy, number_rating, rating) %>%
  filter(country == "USA") %>%
  group_by(candy) %>%
  summarise(USA_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(USA_count)) %>%
  head(5)

```
Answer: any_full_sized_candy_bar with reeses second 

*'UK'*

```{r}
rating_candy_data %>%
  select(country, candy, number_rating, rating) %>%
  filter(country == "UK") %>%
  group_by(candy) %>%
  summarise(UK_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(UK_count)) %>%
  head(5)

```
Answer: any_full_sized_candy_bar with lindt_truffle second

*'CANADA'*

```{r}
rating_candy_data %>%
  select(country, candy, number_rating, rating) %>%
  filter(country == "CANADA") %>%
  group_by(candy) %>%
  summarise(CANADA_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(CANADA_count)) %>%
  head(5)

```
Answer: any_full_sized_candy_bar with kit_kat second

*'The rest of the world"*
```{r}
rating_candy_data %>%
  select(country, candy, number_rating, rating) %>%
  filter(
    country == "EU" |
      country == "ASIA" |
      country == "Other"
  ) %>%
  group_by(candy) %>%
  summarise(other_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(other_count)) %>%
  head(5)

```
Answer: any_full_sized_candy_bar with tolberone second.

<br>