---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
candy_all_data <- read_csv("clean_data/candy_all_data.csv")
```

What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
candy_all_data %>% 
  filter(!is.na(rating)) %>% 
  summarise(rating = n())
```
Answer: 613252



What was the average age of people who are going out trick or treating?

```{r}

candy_all_data %>% 
  summarise(age = mean(age)) %>% 
  round()

```
Answer: rounded 39


What was the average age of people who are not going trick or treating?


```{r}

candy_all_data %>% 
  filter(going_trick_or_treating == "No") %>% 
  summarise(age = mean(age)) %>% 
  round()


```
Answer: rounded 39

For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
candy_all_data %>% 
  select(rating) %>% 
  filter(rating == "JOY") %>% 
  summarise(rating = n())

candy_all_data %>% 
  select(rating) %>% 
  filter(rating == "MEH") %>% 
  summarise(rating = n())

candy_all_data %>% 
  select(rating) %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(rating = n())

```
JOY - 263022
MEH - 76469
DESPAIR = 273761

ANSWER: DESPAIR


How many people rated Starburst as despair?

```{r}
candy_all_data %>% 
  filter(candy == "starburst") %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(candy = n())
```

ANSWER: 1866


For the next three questions, count despair as -1, joy as +1, and meh as 0.

```{r}

rating_candy_data <- candy_all_data %>%
  mutate(number_rating = case_when(
    str_detect(rating, pattern = "DESPAIR") ~ -1,
    str_detect(rating, pattern = "JOY") ~ 1,
    str_detect(rating, pattern = "MEH") ~ 0,
  ))

```



What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
rating_candy_data %>%
  select(your_gender, candy, number_rating, rating) %>%
  filter(your_gender == "Male") %>%
  group_by(candy) %>%
  summarise(male_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(male_count)) %>%
  slice_max(male_count)

rating_candy_data %>%
  select(your_gender, candy, number_rating, rating) %>%
  filter(your_gender == "Female") %>%
  group_by(candy) %>%
  summarise(female_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(female_count)) %>%
  slice_max(female_count)

rating_candy_data %>%
  select(your_gender, candy, number_rating, rating) %>%
  filter(your_gender == "Other") %>%
  group_by(candy) %>%
  summarise(other_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(other_count)) %>%
  slice_max(other_count)

```
male = any_full_sized_candy_bar
female = any_full_sized_candy_bar
Other = any_full_sized_candy_bar


What was the most popular candy bar in each year?


```{r}
rating_candy_data %>%
  select(year, candy, number_rating, rating) %>%
  filter(year == "2015") %>%
  group_by(candy) %>%
  summarise(year_2015_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(year_2015_count)) %>%
  slice_max(year_2015_count)

rating_candy_data %>%
  select(year, candy, number_rating, rating) %>%
  filter(year == "2016") %>%
  group_by(candy) %>%
  summarise(year_2016_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(year_2016_count)) %>%
  slice_max(year_2016_count)

rating_candy_data %>%
  select(year, candy, number_rating, rating) %>%
  filter(year == "2017") %>%
  group_by(candy) %>%
  summarise(year_2017_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(year_2017_count)) %>%
  slice_max(year_2017_count)
```
2015 = any_full_sized_candy_bar
2016 = any_full_sized_candy_bar
2017 = any_full_sized_candy_bar

What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}


rating_candy_data %>%
  select(which_country_do_you_live_in, candy, number_rating, rating) %>%
  filter(which_country_do_you_live_in == "USA") %>%
  group_by(candy) %>%
  summarise(USA_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(USA_count)) %>%
  slice_max(USA_count)

rating_candy_data %>%
  select(which_country_do_you_live_in, candy, number_rating, rating) %>%
  filter(which_country_do_you_live_in == "UK") %>%
  group_by(candy) %>%
  summarise(UK_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(UK_count)) %>%
  slice_max(UK_count)

rating_candy_data %>%
  select(which_country_do_you_live_in, candy, number_rating, rating) %>%
  filter(which_country_do_you_live_in == "CANADA") %>%
  group_by(candy) %>%
  summarise(CANADA_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(CANADA_count)) %>%
  slice_max(CANADA_count)


rating_candy_data %>%
  select(which_country_do_you_live_in, candy, number_rating, rating) %>%
  filter(
    which_country_do_you_live_in == "EU" |
      which_country_do_you_live_in == "ASIA" |
      which_country_do_you_live_in == "Other"
  ) %>%
  group_by(candy) %>%
  summarise(other_count = sum(number_rating, na.rm = TRUE)) %>%
  arrange(desc(other_count)) %>%
  slice_max(other_count)


```
USA - any_full_sized_candy_bar
CANADA - any_full_sized_candy_bar
UK - draw -- lindt_truffle and any_full_sized_candy_bar
Other - any_full_sized_candy_bar