---
title: "TASK:3 SEA BIRDS 🐧 "
output: html_notebook
---

#**INTRO** 

This dataset looks at the sightings of different sea birds, it also includes the ships that had each sighting.


#**CLEANING PROCESS🧹**

For the cleaning proseece I done a couple of things:

Ship_data - 
* This data frame started off with a lot of irrelevant info for the questions we had, these meant I simplified it down to just information I needed.

Bird_data - 
* **Columns:** The Raw data also had a lot information that was not required for our analysis
             
* **Removing from string:** The names of the birds had a lot of stuff in the column that I removed. *'e.g Diomedea impavida / melanophrys AD '* with the AD being removed with following pattern "AD|JUV|IMM|SUBAD|PL[0-9]+"

* **Removing no count:** I also removed any data that didn’t t have a count as to me this was a wasted entry in the data


#**LET'S ANALYSE🔎** 


```{r}
library(tidyverse)
library(here)
```

```{r}
ship_and_bird_clean <- read_csv(here("raw_data/ship_and_bird_clean.csv"))
```

**SIGHTINGS**

**Which bird had the most individual sightings?**

```{r}
ship_and_bird_clean %>% 
  select(species_scientific, species, species_abbreviation, count) %>% 
  filter(count == 1) %>% 
  group_by(species, species_scientific, species_abbreviation) %>% 
  summarise(sightings = sum(count, na.rm = TRUE), .groups = "drop") %>% 
  arrange(desc(sightings)) %>% 
  head(1)
```
Answer: Wandering albatross sensu lato with 1707 

**HIGHEST COUNT**

**Which bird had the highest total count?**

```{r}
ship_and_bird_clean %>% 
  select(species_scientific, species, species_abbreviation, count) %>% 
  group_by(species, species_scientific, species_abbreviation) %>% 
  summarise(sightings = sum(count, na.rm = TRUE), .groups = "drop") %>% 
  arrange(desc(sightings)) %>% 
  head(1)
```
Answer: Short tailed shearwater

**LATITUDE**

**Which bird had the highest total count above a latitude of -30?**

```{r}
ship_and_bird_clean %>% 
  filter(lat > -30) %>% 
  group_by(species, species_scientific, species_abbreviation) %>% 
  summarise(sightings = sum(count, na.rm = TRUE), .groups = "drop") %>% 
  arrange(desc(sightings)) %>% 
  head(1)
  
```
Answer: Wedge tailed shearwater

**LONLEY BIRDS**

**How many different types of birds were only ever seen in groups of 1?**

```{r}

ship_and_bird_clean %>% 
  group_by(species, species_scientific, species_abbreviation) %>% 
  summarise(mean_sightins = mean(count, na.rm = TRUE), .groups = "drop") %>% 
  filter(mean_sightins == 1) %>% 
  nrow()

```
Answer: 37

**PENGUINS**

**How many penguins were seen?**

```{r}
ship_and_bird_clean %>% 
  mutate(is_penguin = str_detect(species, pattern = "penguin")) %>% 
  filter(is_penguin == TRUE) %>% 
  summarise(sum(count, na.rm = TRUE)) 

```
Answer: 154
